---
title: "Coastal Units with hyRefactor"
author: "dblodgett@usgs.gov"
output: rmarkdown::html_vignette
always_allow_html: true
vignette: >
  %\VignetteIndexEntry{coastal_units}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
```{r setup, include = FALSE}
  knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width=6,
  fig.height=4
)
options("rgdal_show_exportToProj4_warnings"="none",
        rmarkdown.html_vignette.check_title = FALSE)
library(magrittr)

fdr_path <- "../../gfv2/workspace/data/fdrfac/NHDPlusCA/NHDPlus18/NHDPlusFdrFac18c/fdr/"
fac_path <- "../../gfv2/workspace/data/fdrfac/NHDPlusCA/NHDPlus18/NHDPlusFdrFac18c/fac/"
```

This article demonstrates how to work with coastal catchments and small coastal basins in an hyRefactor workflow.

First, we use `nhdplusTools` to download and plot some data along the California coast.

```{r, plot}
library(hyRefactor)

temp_gpkg <- tempfile(fileext = ".gpkg")

bbox <- sf::st_bbox(c(xmin = -124.4, ymin = 39.8, xmax = -123.7, ymax = 40.3), 
                    crs = sf::st_crs(4326))
nhd_data <- nhdplusTools::plot_nhdplus(bbox = bbox, gpkg = temp_gpkg, overwrite = TRUE)
```

Now that we have some data to work with, we need to subset it so we have a few complete basins to work with. 

_Note that both basins that terminate at the coast AND the coastal flowlines are included in out subset below._

First we use the `nhdplusTools::navigate_nldi` for each terminal flowline in our subset to find complete basins then we use the `nhdplusTools::subset_nhdplus` to get the data we need.
```{r}
terminals <- dplyr::filter(nhd_data$flowline, TerminalFl == 1)

coastal <- dplyr::filter(nhd_data$flowline, FTYPE == "Coastline")

flowline <- dplyr::bind_rows(
  lapply(terminals$COMID, function(x) {
    nhdplusTools::navigate_nldi(list(featureSource = "comid", featureID = x), 
                                mode = "UT", distance_km = 999)$UT_flowlines
}))

sub <- lapply(nhdplusTools::subset_nhdplus(c(as.integer(flowline$nhdplus_comid), coastal$COMID), 
                                    nhdplus_data = "download", flowline_only = FALSE),
              nhdplusTools::align_nhdplus_names)


mapview::mapview(sub$NHDFlowline_Network) + 
  mapview::mapview(sub$CatchmentSP, color = "darkgrey", col.regions = "tan")
```
Now that we have our subset ready to go, we'll need to get some flow direction and flow accumulation data to use with hyRefactor later on.
```{r fdr_fac, eval=FALSE}
fdr_dir = file.path(tempdir(check = TRUE), "fdr_fac")
dir.create(fdr_dir, showWarnings = FALSE)

if(!dir.exists(file.path(fdr_dir, "NHDPlusCA/"))) {
  hyRefactor::download_fdr_fac(fdr_dir, regions = "18")
}

fdr_path <- file.path(fdr_dir, "NHDPlusCA/NHDPlus18/NHDPlusFdrFac18c/fdr/")
fac_path <- file.path(fdr_dir, "NHDPlusCA/NHDPlus18/NHDPlusFdrFac18c/fac/")
```
```{r}
library(terra)
fdr <- terra::rast(fdr_path)
fac <- terra::rast(fac_path)
crs <- terra::crs(fac)
fdr <- terra::crop(fdr, sf::st_transform(sub$CatchmentSP, crs))
fac <- terra::crop(fac, sf::st_transform(sub$CatchmentSP, crs))
plot(fdr)
```
Now that we have all our data ready to go, we can get our networks that we want to refactor pulled out from our coastal catchments and small coastal basins. 
```{r}
net <- sub$NHDFlowline_Network
cats <- sub$CatchmentSP

coastal <- net[net$FTYPE == "Coastline", ]
coastal_cats <- cats[cats$FEATUREID %in% coastal$COMID, ]

net <- dplyr::filter(net, !COMID %in% coastal$COMID)
cats <- cats[cats$FEATUREID %in% net$COMID, ]
                       
nhd_outlets <- dplyr::filter(net, TerminalFl == 1)

sf::st_geometry(nhd_outlets) <- sf::st_geometry(nhdplusTools::get_node(nhd_outlets, "end"))

mapview::mapview(net) + 
  mapview::mapview(cats, color = "darkgrey", col.regions = "tan") + 
  mapview::mapview(coastal, color = "lightblue") + 
  mapview::mapview(coastal_cats, color = "darkgrey", col.regions = "brown") + 
  mapview::mapview(nhd_outlets, color = "black")
```

Now we will identify coastal basins that we want to lump into coastal catchments as part of the refactor workflow.
```{r}
min_da_km <- 10

little_terminal <- dplyr::filter(net, TerminalPa %in% 
                                   dplyr::filter(nhd_outlets, 
                                                 TotDASqKM <= min_da_km & 
                                                   TerminalFl == 1)$TerminalPa)

outlets <- dplyr::select(nhd_outlets, COMID) %>%
  dplyr::mutate(type = "terminal") %>%
  dplyr::filter(COMID %in% cats$FEATUREID) %>%
  dplyr::mutate(keep = ifelse(COMID %in% little_terminal$COMID, "temporary", "keep"))

mapview::mapview(outlets, zcol = "keep") + 
  mapview::mapview(net)
```

We can now run `refactor_nhdplus` and `reconcile_catchment_divides`. 

_Note that the `exclude_cats` parameter is set to all outlet flowlines in small basins that were identified above._

_Also note that the `net` variable no longer contains coastal flowlines._
```{r}
tf <- file.path(tempfile(fileext = "tf.gpkg"))
tr <- file.path(tempfile(fileext = "tr.gpkg"))

refactor_nhdplus(nhdplus_flines = net, 
                 split_flines_meters = 100000, 
                 split_flines_cores = 1, 
                 collapse_flines_meters = 2000,
                 collapse_flines_main_meters = 2000,
                 out_refactored = tf, 
                 out_reconciled = tr, 
                 three_pass = TRUE, 
                 purge_non_dendritic = FALSE, 
                 exclude_cats = unique(c(outlets$COMID, little_terminal$COMID)),
                 warn = FALSE)

refactored <- sf::st_transform(sf::read_sf(tf), crs)
reconciled <- sf::st_transform(sf::read_sf(tr), crs)
cats <- sf::st_transform(cats, crs)
sf::st_precision(cats) <- 10

divides <- reconcile_catchment_divides(catchment = cats,
                                       fline_ref = refactored,
                                       fline_rec = reconciled,
                                       fdr = fdr_path,
                                       fac = fac_path,
                                       para = 1) 

mapview::mapview(reconciled) + 
  mapview::mapview(divides)
```
Finally, we can identify the outlets 
```{r}
keep_outlets <- dplyr::filter(outlets, keep == "keep") %>%
  dplyr::select(COMID, type)

mapped_outlets <- map_outlet_ids(keep_outlets, reconciled) %>%
  dplyr::filter(COMID %in% keep_outlets$COMID)

zero_order <- list(basin = little_terminal$COMID, zero = coastal$COMID)

agg_cats <- aggregate_catchments(flowpath = reconciled, 
                                 divide = divides, 
                                 outlets = dplyr::select(mapped_outlets, ID, type),
                                 zero_order = zero_order,
                                 coastal_cats = sf::st_transform(coastal_cats, sf::st_crs(divides)),
                                 da_thresh = 1, 
                                 only_larger = TRUE)
mapview::mapview(agg_cats)
```

