---
title: "Refactor Catchments"
author: "dblodgett@usgs.gov"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{refactor_catchment}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width=6,
  fig.height=4
)
options(rmarkdown.html_vignette.check_title = FALSE)
options(scipen = 9999)
```
# Refactoring Catchments with NHDPlusTools

## Load Network and Refactor

For this example, we will use some data available from the nhdplusTools package for testing and examples. `fac_sample`, `fdr_sample`, `flowline_sample`, `catchment_sample` are all included as sample data in `nhdplusTools`.

Before we can do anything with the catchments, we need to run the NHDPlus Refactor workflow documented over in `vignette("refactor_nhdplus")`.

```{r refactor}
library(sf)
library(nhdplusTools)
library(hyRefactor)
library(raster)

source(system.file("extdata", "walker_data.R", package = "hyRefactor"))

ref <- tempfile(fileext = ".gpkg")
rec <- tempfile(fileext = ".gpkg")

refactor_nhdplus(nhdplus_flines = walker_flowline,
                 split_flines_meters = 2000,
                 collapse_flines_meters = 1000,
                 collapse_flines_main_meters = 1000,
                 split_flines_cores = 2,
                 out_refactored = ref,
                 out_reconciled = rec,
                 three_pass = TRUE,
                 purge_non_dendritic = FALSE,
                 warn = FALSE)

flowline_ref <- read_sf(ref)
flowline_rec <- read_sf(rec)
```

Let's start by looking at a single sample catchment. In the first plot, you can see the input is a single polygon and a set of catchment flowlines. For reference, the Flow Direction Raster (FDR) and Flow Accumulation Raster (FAC) are also plotted.

```{r single_plots}
sample_catchment <- dplyr::filter(walker_catchment, FEATUREID == 5329435)

sample_flowline <- dplyr::filter(flowline_ref, as.integer(COMID) == 5329435)

plot(st_geometry(sample_catchment))
plot(sample_flowline["COMID"], add = TRUE)

raster::plot(walker_fdr)
raster::plot( walker_fac)
```

Now we can run the `split_catchment_divide()` function which is designed to take one catchment and its associated (split) flowlines. The split flowlines are the "refactored" but not "reconciled" output of the `nhdplus_refactor()` function we ran above. Here we run the `split_catchment()` function and plot up the resulting data to show what it did.

```{r single_split}
split_cat <- split_catchment_divide(sample_catchment, sample_flowline, walker_fdr,  walker_fac)

plot(st_geometry(split_cat), col = NA, border = "red")
plot(st_geometry(sample_catchment), add = TRUE)
plot(sample_flowline["COMID"], lwd = 2, add= TRUE)
```

As you can see, the flowline in question was split into five pieces by `nhdplus_refactor()` and the cooresponding catchment (black) was similarly broken up into 5 sub-catchments (red).

The `split_catchment_divide()` function can be run against a set of catchments using the `reconcile_catchment_divides()` function. This functiona can call `split_catchment()` in parallel, unions catchments according to the output of `nhdplus_refactor()`, and assembles the results back into an `sf` `data.frame`.

```{r cat_set}
split_cats <- reconcile_catchment_divides(catchment = walker_catchment, 
                                          fline_ref = flowline_ref, 
                                          fline_rec = flowline_rec, 
                                          fdr = walker_fdr, 
                                          fac = walker_fac, 
                                          para = 4)

plot(st_geometry(split_cats), col = NA, border = "red")
plot(st_geometry(walker_catchment), col = NA, border = "black")
plot(st_geometry(split_cats), col = NA, border = "red")
plot(st_geometry(walker_catchment), col = NA, border = "black", add = TRUE)
```

## Aggregation

With our flowline network and catchments all refactored and the geometry reconciled, we could also aggregate the resulting network to a selected set of outlet locations. 

```{r aggregate}
outlets <- data.frame(ID = c(31, 3, 5, 1, 45, 92),
                      type = c("outlet", "outlet", "outlet", "terminal", "outlet", "outlet"),
                      stringsAsFactors = FALSE)

aggregated <- aggregate_catchments(flowpath = flowline_rec, 
                                   divide = split_cats,
                                   outlets = outlets)

plot(aggregated$cat_sets$geom, lwd = 3, border = "red")
plot(split_cats$geom, lwd = 1.5, border = "green", col = NA, add = TRUE)
plot(walker_catchment$geom, lwd = 1, add = TRUE)
plot(walker_flowline$geom, lwd = .7, col = "blue", add = TRUE)
#'
plot(aggregated$cat_sets$geom, lwd = 3, border = "black")
plot(aggregated$fline_sets$geom, lwd = 3, col = "red", add = TRUE)
plot(walker_flowline$geom, lwd = .7, col = "blue", add = TRUE)
```

## Identifiers

So you want to know how the identifiers from the original NHDPlus Catchments relate to the output? Let's look at what's in the output. 

```{r identifiers} 
(input_ids <- walker_flowline$COMID)

(refactored_ids <- flowline_rec$ID)

(refactored_id_mapping <- flowline_rec$member_COMID)

(aggregated_ids <- aggregated$cat_sets$ID)

(aggregated_id_mapping <- aggregated$cat_sets$set)
```

This may look like a complicated mess but there's structure here. Stepping through the identifiers, we have:

1. NHDPlusV2 COMIDs (shared between catchment divides and flowlines)
1. Refactored Catchment Identifiers (shared between catchment divides and flowpaths)

There are two caveats here.

1. Because the source NHDPlusV2 catchments were split, the relationship between refactored catchments and source catchments requires a sequence (.1, .2, .. .10, etc. -- upstream to downstream) to differentiate the parts.
2. Aggregated catchments use the identifer of the outlet catchment.

Given these caveats, we can build a complete lookup table from source catchment to output aggregate identifier.

```{r lookup}

refactor_lookup <- dplyr::select(st_drop_geometry(flowline_rec), ID, member_COMID) %>%
  dplyr::mutate(member_COMID = strsplit(member_COMID, ",")) %>%
  tidyr::unnest(cols = member_COMID) %>%
  dplyr::mutate(NHDPlusV2_COMID = as.integer(member_COMID)) %>% # note as.integer truncates
  dplyr::rename(reconciled_ID = ID)

aggregate_lookup_fline <- dplyr::select(st_drop_geometry(aggregated$fline_sets), ID, set) %>%
  tidyr::unnest(cols = set) %>%
  dplyr::rename(aggregated_flowline_ID = ID, reconciled_ID = set)

aggregate_lookup_catchment <- dplyr::select(st_drop_geometry(aggregated$cat_sets), ID, set) %>%
  tidyr::unnest(cols = set) %>%
  dplyr::rename(aggregated_catchment_ID = ID, reconciled_ID = set)

(lookup_table <- tibble::tibble(NHDPlusV2_COMID = input_ids) %>%
  dplyr::left_join(refactor_lookup, by = "NHDPlusV2_COMID") %>%
  dplyr::left_join(aggregate_lookup_fline, by = "reconciled_ID") %>%
  dplyr::left_join(aggregate_lookup_catchment, by = "reconciled_ID"))

```

Ta Da!! Have fun and [report bugs here](https://github.com/dblodgett-usgs/nhdplusTools/issues). 